---
title: "Celebrity COVID Cases"
date: "`r format(Sys.time(), '%B %d, %Y')`"
urlcolor: blue
header-includes:
  - \usepackage{amsmath}
output:
  html_document:
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, message=FALSE, echo=FALSE}

library(tidyverse) # package that helps "wrangle" data
library(ggplot2); theme_set(theme_bw())
library(magrittr)  # package that lets you use "piping" i.e., %>%
library(knitr)     # package that helps make html and pdfs from r code
library(table1)    # package to make nice tables
library(lattice)
library(GGally)

opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
options(scipen = 999)
```

# Summary of data

```{r readData}
# read in your data. 
# don't need last columns for now. 
data <- read.csv('data/celebrity.csv', na.strings = c(""))[, 1:19] 

# fix some columns for easier analysis
data %<>%
  # change these variables to lowercase
  mutate_at(vars(Test_results, Symptoms_binary), tolower) %>%
  mutate(Category_career=str_trim(Category_career)) %>%
  # change these date variables to character first, weird quirk of R
  mutate_at(vars(contains("Date")), as.character) %>%
  # then change them to "Date" variable
  mutate_at(vars(contains("Date")), as.Date, format="%Y-%m-%d")

```

```{r summarizeData}
# Making new variables that are categorical variables  
data %<>%
  mutate(Age_cat=cut(data$Age, c(15, 25, 35, 55, 85))) %>%
  mutate(Net_worth_cat=cut(data$Net_worth, c(0, 1e6, 1e7, 
                                            1e8, 1e10))) %>%
  mutate(Instagram_followers_cat=cut(data$Instagram_followers, 
                                    c(0, 1.5e5, 5e5, 
                                            5e6, 5e7)))

# Giving our data nice labels for the table below
label(data$Age_cat) <- "Age"
label(data$Category_career) <- "Career"
label(data$Test_country) <- "Country tested"
label(data$Net_worth_cat) <- "Net Worth"
label(data$Instagram_followers_cat) <- "Instagram followers"

# A nice summary table of the celebrites
table1(~Category_career + Age_cat|Sex, data = data)
```

Our current data contain more than three times as many male celebrities as female celebrities. Part of the difference is due to a large number of male athletes being tested and a large difference in the number of male politicians tested. 

The ages of celebrities tested varied from 19 to 81. Most celebrities were 35-55 years old.

```{r}
# Summary tables of the celebrity predictors
data %>%
  filter(!is.na(Net_worth)) %>%
  group_by(Category_career) %>%
  summarise(Mean = mean(Net_worth/1e6, na.rm = TRUE), 
           SD = sd(Net_worth/1e6, na.rm = TRUE), 
           n = n()) %>% 
  kable(digits = 1, caption="Net worth of celebrites tested for COVID (in millions). ")

data %>%
  filter(!is.na(Instagram_followers)) %>%
  group_by(Category_career) %>%
  summarise(Mean = mean(Instagram_followers/1e3, na.rm = TRUE), 
           SD = sd(Instagram_followers/1e3, na.rm = TRUE), 
           n = n()) %>% kable(digits = 1, caption="Instagram followers of celebrites tested for COVID (in thousands). ")
```

Celebrities in the entertainment industry had the highest net worth on average, followed by politicians. Entertainment celebrities also had the most instagram followers followed by celebrities, on average. 

```{r continuousHistograms, fig.cap="Celebrity net worth and followers by career type. "}

data %>%
  ggplot(aes(x=Category_career, y=Net_worth/1e6)) + 
  geom_boxplot() + scale_y_log10()  -> p1

data %>%
  ggplot(aes(x=Category_career, y=Instagram_followers/1e3)) + 
  geom_boxplot() + scale_y_log10()  -> p2

gridExtra::grid.arrange(p1, p2, ncol=1)

```


```{r, fig.cap="Celebrity coronavirus cases by reporting date in our dataset. "}
# Try making an epidemic curve for the celebrities using date of symptoms or date of reporting. 

data %>%
  group_by(Date_reported) %>%
  tally(name="Cases") %>%     # check out the data up to here
  # figure out how to get date of reporting on the x axis
  ggplot(aes(x=Date_reported, y=Cases)) + 
  geom_line() + 
  geom_point() + 
  labs(x="Date of reporting", y="Celebrity cases")

```

Reports of celebrity coronavirus cases peaked on March 19th and has trailed off. 

# Univariate analyses of delay distributions

Here, we want to see the overall delay distributions of all celebrities. 

```{r delayDistributions, fig.cap="Delay distributions of celebrity cases.  "}
data %<>%
  mutate(Exposure_to_reporting=Date_reported-Date_exposure) %>%
  mutate(Symptoms_to_testing=Date_of_Test-Date_symptoms) %>%
  mutate(Testing_to_reporting=Date_reported-Date_of_Test) %>%
  mutate(Symptoms_to_reporting=Date_reported-Date_symptoms) 

intervalData <- data %>%
  select(contains("to_"))  %>%
  mutate_each(as.numeric) %>%
  gather("Interval", "Days")

# this is an easy to use function from the lattice package
densityplot(~Days|Interval, data = intervalData)

data%>%
  select(contains("to_"))  %>%
  mutate_each(as.numeric) %>%
  summary()
```

Celebrities are being tested 2.9 days after symptom onset, on average with a median of 3 days. They are getting their results back 2.4 days after testing, with a median of 1 day. 

# Multivariate analyses

Response variables: 

- exposure to symptoms
- exposure to testing
- exposure to reporting
- symptoms to testing
- symptoms to reporting
- testing to reporting

Predictor variables: 

- age
- sex
- symptoms binary
- career category
- net worth
- insta followers
- country

```{r symptomToTestingContinuous, fig.cap="Symptom to testing delay by age, net worth, and instagram followers.  "}

data %>%
  select(Age, Net_worth, 
         Instagram_followers, 
         Symptoms_to_testing) %>%
  mutate(Net_worth=Net_worth/1e6,
         Instagram_followers=Instagram_followers/1e3) %>% 
  mutate(Symptoms_to_testing=as.numeric(Symptoms_to_testing)) %>%
  mutate_at(vars(Net_worth, Instagram_followers), log) %>%
  ggpairs() + 
  theme(axis.text.x = element_text(angle = 90))
```

Older age is POSITIVELY correlated with symptom to testing delay. Not what we expected. Net worth and instagram followership is not associated with symptom to testing delay. 

```{r testingToReportingContinuous, fig.cap="Testing to reporting delay by age, net worth, and instagram followers.  "}
data %>%
  select(Age, Net_worth, 
         Instagram_followers, 
         Testing_to_reporting) %>%
  mutate(Net_worth=Net_worth/1e6,
         Instagram_followers=Instagram_followers/1e3) %>% 
  mutate(Testing_to_reporting=as.numeric(Testing_to_reporting)) %>%
  mutate_at(vars(Net_worth, Instagram_followers), log) %>%
  ggpairs() + 
  theme(axis.text.x = element_text(angle = 90))
```

There is a slight negative correlation between age and testing to reporting delay. Net worth and instagram followers did not have a relationship with testing to reporting delay. 

```{r symptomsToTestingDiscrete, fig.cap="Symptom to testing delay by sex, career category, and test country. "}

data %>%
  select(Sex, Symptoms_to_testing) %>%
  filter(!is.na(Sex)) %>%
  mutate(Symptoms_to_testing=as.numeric(Symptoms_to_testing)) %>%
  ggplot(aes(x=Sex, y=Symptoms_to_testing)) + 
  geom_boxplot() -> p1

data %>%
  select(Category_career, Symptoms_to_testing) %>%
  filter(Category_career%in%c("Athlete", "Politician", 
                              "Entertainment")) %>%
  mutate(Symptoms_to_testing=as.numeric(Symptoms_to_testing)) %>%
  ggplot(aes(x=Category_career, y=Symptoms_to_testing)) + 
  geom_boxplot() -> p2

data %>%
  select(Test_country, Symptoms_to_reporting) %>%
  filter(!is.na(Test_country)) %>%
  mutate(Test_country=ifelse(Test_country=="USA", "USA", "Other")) %>%
  mutate(Symptoms_to_reporting=as.numeric(Symptoms_to_reporting)) %>%
  ggplot(aes(x=Test_country, y=Symptoms_to_reporting)) + 
  geom_boxplot() -> p3

gridExtra::grid.arrange(p1, p2, p3) 
```

The most apparent difference in symptom to testing delay is by celebrity career type. Athletes were tested very quickly following a scare with a few infected players lowering their average. 

```{r testingToReportingDiscrete, fig.cap="Testing to reporting delay by sex, career category, and test country. "}

data %>%
  select(Sex, Testing_to_reporting) %>%
  filter(!is.na(Sex)) %>%
  mutate(Testing_to_reporting=as.numeric(Testing_to_reporting)) %>%
  ggplot(aes(x=Sex, y=Testing_to_reporting)) + 
  geom_boxplot() -> p1

data %>%
  select(Category_career, Testing_to_reporting) %>%
  filter(Category_career%in%c("Athlete", "Politician", 
                              "Entertainment")) %>%
  mutate(Testing_to_reporting=as.numeric(Testing_to_reporting)) %>%
  ggplot(aes(x=Category_career, y=Testing_to_reporting)) + 
  geom_boxplot() -> p2

data %>%
  select(Test_country, Testing_to_reporting) %>%
  filter(!is.na(Test_country)) %>%
  mutate(Test_country=ifelse(Test_country=="USA", "USA", "Other")) %>%
  mutate(Testing_to_reporting=as.numeric(Testing_to_reporting)) %>%
  ggplot(aes(x=Test_country, y=Testing_to_reporting)) + 
  geom_boxplot() -> p3

gridExtra::grid.arrange(p1, p2, p3) 
```

The most interesting difference here is that even celebrities in the US are (sometimes) having to wait 5-8 days for test results. But on average, they are getting their test result back in 1 day, like the rest of the world. 


